---
title: "Methods for Data Science - Final Project"
author: "Chaitya Mehta and Pruthvi Patel"
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(dplyr)
library(RCurl)
library(funModeling) 
library(tidyverse) 
library(Hmisc)
#importing data from Github
url1 = 'https://raw.githubusercontent.com/GitPruthvi/TelecomChurnRatePredictor/master/customer_data.csv'
customerData = read.csv(url(url1))

url2 = 'https://raw.githubusercontent.com/GitPruthvi/TelecomChurnRatePredictor/master/churn_data.csv'
churnData = read.csv(url(url2))

url3 = 'https://raw.githubusercontent.com/GitPruthvi/TelecomChurnRatePredictor/master/internet_data.csv'
internetData = read.csv(url(url3))

#merging 3 dataframes into 1
mainData = merge(merge(customerData,churnData,by="customerID"),internetData, by="customerID")

head(mainData,5)
glimpse(mainData)
dim(mainData)
summary(mainData)
mainData$customerID=NULL #removing customerID from the dataset as it is just the primary key to identify the customer, it doesn't hold any significance
```

# EDA

```{r}
#install.packages("plotrix")
library(plotrix)
length(which(mainData$Churn == "No"))
length(which(mainData$Churn == "Yes"))
slices <- c(5174,1869)
lbls <- c("Not Churn", "Churn")
pct <- (slices/sum(slices)*100)
pct = sprintf("%0.2f", round(pct, digits = 2))
lbls <- paste(lbls,"-", pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie3D(slices,labels=lbls,main="Churn vs Not Churn",col=rainbow(length(slices)),radius = 2.5)

```
```{r}
#install.packages("plotrix")
library(plotrix)
length(which(mainData$gender == "Male"))
length(which(mainData$gender == "Female"))
slices <- c(3555,3488)
lbls <- c("Male","Female")
pct <- (slices/sum(slices)*100)
pct = sprintf("%0.2f", round(pct, digits = 2))
lbls <- paste(lbls,"\n", pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie3D(slices,labels=lbls,main="Male vs Female",col=rainbow(length(slices)))

```

```{r}
library(dplyr)
M_C=filter(mainData, (mainData$gender=="Male" & mainData$Churn=="Yes"))
M_NC=filter(mainData, (mainData$gender=="Male" & mainData$Churn=="No"))
nrow(M_C)
nrow(M_NC)
# Number of Male who left the company (churned i.e. M_C) are 930
# Number of Male who did not left the company (not churned i.e. M_NC) are 2625

F_C=filter(mainData, (mainData$gender=="Female" & mainData$Churn=="Yes"))
F_NC=filter(mainData, (mainData$gender=="Female" & mainData$Churn=="No"))
nrow(F_C)
nrow(F_NC)
# Number of Female who left the company (churned i.e. F_C) are 939
# Number of Female who did not left the company (not churned i.e. F_NC) are 2549

slices1 <- c(930,2625)
lbls <- c("Yes","No")
pct <- (slices1/sum(slices1)*100)
pct = sprintf("%0.2f", round(pct, digits = 2))
lbls <- paste(lbls,"-", pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie3D(slices1,labels=lbls,main="Male Customers - Not Churn vs Churn",col=rainbow(length(slices1)))

slices1 <- c(939,2549)
lbls <- c("Yes","No")
pct <- (slices1/sum(slices1)*100)
pct = sprintf("%0.2f", round(pct, digits = 2))
lbls <- paste(lbls,"-", pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie3D(slices1,labels=lbls,main="Female Customers - Churn vs Not Churn",col=rainbow(length(slices1)))


```


```{r}
counts <- table(mainData$gender, mainData$Churn)
# get percentages of vars

pcnts <- scale(counts, FALSE, colSums(counts))*100
# plot barplot

bp <- barplot(pcnts, beside=TRUE, col=c("lightblue","red"),xlab="Gender - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space=c(0,5))
legend("center", legend=c("Female", "Male"), bty="n", fill=c("lightblue", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```
```{r}
counts <- table(mainData$SeniorCitizen, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red"),xlab="SeniorCitizen - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("Not a SeniorCitizen", "SeniorCitizen"), bty="n", fill=c("lightblue", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
#colnames(mainData)
counts <- table(mainData$Partner, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red"),xlab="Partner - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("Doesn't have a Partner", "Have a Partner"), bty="n", fill=c("lightblue", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```
```{r}
#colnames(mainData)
counts <- table(mainData$Dependents, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red"),xlab="Dependents - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space=c(0,5))
legend("center", legend=c("No Dependents", "Have Dependents"), bty="n", fill=c("lightblue", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```
```{r}
#colnames(mainData)
counts <- table(mainData$PhoneService, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red"),xlab="PhoneService - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space=c(0,5))
legend("center", legend=c("Doesn't uses PhoneService", "Uses PhoneService"), bty="n", fill=c("lightblue", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```
```{r}
colnames(mainData)
counts <- table(mainData$MultipleLines, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "green","red"),xlab="MultipleLines - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No","No phone Service ", "Yes"), bty="n", fill=c("lightblue","green", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```
```{r}
colnames(mainData)
counts <- table(mainData$InternetService, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "green","red"),xlab="InternetService - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("Fiber optic","No", "DSL"), bty="n", fill=c("lightblue","green", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```
```{r}
#colnames(mainData)
counts <- table(mainData$OnlineSecurity, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue","red","green"),xlab="OnlineSecurity - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No", "No internet service", "Yes"), bty="n", fill=c("lightblue", "red","green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
#colnames(mainData)
counts <- table(mainData$OnlineBackup, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue","red","green"),xlab="OnlineBackup - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No", "No internet service", "Yes"), bty="n", fill=c("lightblue", "red","green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
#colnames(mainData)
counts <- table(mainData$DeviceProtection, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red","green"),xlab="DeviceProtection - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No", "No internet service", "Yes"), bty="n", fill=c("lightblue", "red","green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
counts <- table(mainData$TechSupport, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red", "green"),xlab="TechSupport - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No", "No internet service", "Yes"), bty="n", fill=c("lightblue", "red", "green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
counts <- table(mainData$StreamingTV, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red", "green"),xlab="StreamingTV - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No", "No internet service", "Yes"), bty="n", fill=c("lightblue", "red", "green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
counts <- table(mainData$StreamingMovies, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red", "green"),xlab="StreamingMovies - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No", "No internet service", "Yes"), bty="n", fill=c("lightblue", "red", "green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
counts <- table(mainData$Contract, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue","red", "green"),xlab="Contract - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("Month-to-Month","Two Year", "One Year"), bty="n", fill=c("lightblue","red", "green"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
counts <- table(mainData$PaperlessBilling, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue", "red"),xlab="PaperlessBilling - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("No","Yes"), bty="n", fill=c("lightblue", "red"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```

```{r}
counts <- table(mainData$PaymentMethod, mainData$Churn)

# get percentages of vars
pcnts <- scale(counts, FALSE, colSums(counts))*100

# plot barplot
bp <- barplot(pcnts, beside=TRUE, col=c("lightblue","green", "red","orange"),xlab="PaymentMethod - Non Churn vs Churn", ylab="Frequency (%)", border=NA,space = c(0,5))
legend("center", legend=c("Electronic Check","Bank Transfer", "Mailed check","Credit Card"), bty="n", fill=c("lightblue","green", "red","orange"), border=NA)
text(bp, 1, round(pcnts, 2), cex=1, pos=3, col=c("black"))
```



```{r}
#install.packages("dplyr")
library(dplyr)
churn.mainData = filter(mainData, Churn=="Yes")
notchurn.mainData = filter(mainData, Churn=="No")
#install.packages("DataExplorer")
library(DataExplorer)
par(mfrow = c(2, 2))
plot_density(churn.mainData, title = "        MonthlyCharges, Tenure, TotalCharges - Customer Churn data")
plot_histogram(churn.mainData, title = "        MonthlyCharges, Tenure, TotalCharges - Customer Churn data")
plot_density(notchurn.mainData, title = "        MonthlyCharges, Tenure, TotalCharges - Customer non Churn data")
plot_histogram(notchurn.mainData, title = "        MonthlyCharges, Tenure, TotalCharges - Customer non Churn data")
```




# Data Cleaning
```{r}
library(funModeling) 
#copying the dataset into another which needs to be cleaned
cleanData = data.frame(mainData)
#checking for NA or blank values in the dataset
status(cleanData)
#11 rows with NA for TotalCharges variable

#confirming  the number of NA rows for TotalCharges variable
table(is.na(cleanData$TotalCharges))

#accessing the data for all the 11 rows which has NA values for TotalCharges variable
cleanData[is.na(cleanData$TotalCharges), ]

#taking a not of one customerID who had NA value so that we can confirm later to see if the value of TotalCharges column has been replaced by the mean value  or not "1371-DWPAZ"

#Replacing NA values in TotalCharges column with the mean value of the entire column
cleanData$TotalCharges <- ifelse(is.na(cleanData$TotalCharges), median(cleanData$TotalCharges, na.rm=TRUE), cleanData$TotalCharges) 

#checking the median value of the entire column
median(cleanData$TotalCharges)

#confirming that there are no NA values in the column
table(is.na(cleanData$TotalCharges))

#replacing special  characters like  "-", "(", ")" with a blank space from data values of the dataset 
cleanData$Contract=gsub("-", "", cleanData$Contract)
cleanData$PaymentMethod=gsub("[()]", "", cleanData$PaymentMethod)
cleanData$PaymentMethod=gsub("[()]", "", cleanData$PaymentMethod)

#replacing the blank space with . in the datset as later we will be creating dummy variables and for that we don't want these blank space in the column name
cleanData$Contract<-str_replace_all(cleanData$Contract, c(" " = "." , "," = "" ))
cleanData$PaymentMethod<-str_replace_all(cleanData$PaymentMethod, c(" " = "." , "," = "" ))
cleanData$InternetService<-str_replace_all(cleanData$InternetService, c(" " = "." , "," = "" ))

head(cleanData,5)
glimpse(cleanData)
dim(cleanData)
summary(cleanData)

```



```{r}
#changing all the values from True/Fales and Yes/No to 1 and 0 
cleanData$Churn <- ifelse(cleanData$Churn == "Yes", 1, 0) # 1=Yes / 0=No
cleanData$gender <- ifelse(cleanData$gender == "Male", 1, 0) # 1=male / 0=female
#Senior citizen already in 0 and 1
cleanData$Partner <- ifelse(cleanData$Partner == "Yes", 1, 0) # 1=Yes / 0=No
cleanData$Dependents <- ifelse(cleanData$Dependents == "Yes", 1, 0) # 1=Yes / 0=No
#Tenure in numeric form
cleanData$PhoneService <- ifelse(cleanData$PhoneService == "Yes", 1, 0) # 1=Yes / 0=No

#cleanData$Contract <- ifelse(cleanData$Contract == "Yes", 1, 0) # One year / Two year / Month-to-month
cleanData$Contract = as.factor(cleanData$Contract)
cleanData$PaperlessBilling <- ifelse(cleanData$PaperlessBilling == "Yes", 1, 0) # 1=Yes / 0=No

#cleanData$PaymentMethod <- ifelse(cleanData$PaymentMethod == "Yes", 1, 0) # (Mailed check) / (Electronic check) / (Credit card (automatic)) / (Bank transfer (automatic))
cleanData$PaymentMethod = as.factor(cleanData$PaymentMethod)
# Monthly charges in numeric form
# Total charges in numeric form

cleanData$MultipleLines <- ifelse(cleanData$MultipleLines == "Yes", 1, 0) # 1=Yes / 0=No, No phone services

#cleanData$InternetService <- ifelse(cleanData$InternetService == "Yes", 1, 0) # DSL / Fiber optic	/ No
cleanData$InternetService = as.factor(cleanData$InternetService)

cleanData$OnlineSecurity <- ifelse(cleanData$OnlineSecurity == "Yes", 1, 0) # 1=Yes / 0=No, No phone services
cleanData$OnlineBackup <- ifelse(cleanData$OnlineBackup == "Yes", 1, 0) # 1=Yes / 0=No, No phone services
cleanData$DeviceProtection  <- ifelse(cleanData$DeviceProtection  == "Yes", 1, 0) # 1=Yes / 0=No, No phone services
cleanData$TechSupport <- ifelse(cleanData$TechSupport == "Yes", 1, 0) # 1=Yes / 0=No, No phone services
cleanData$StreamingTV <- ifelse(cleanData$StreamingTV == "Yes", 1, 0) # 1=Yes / 0=No, No phone services
cleanData$StreamingMovies <- ifelse(cleanData$StreamingMovies == "Yes", 1, 0) # 1=Yes / 0=No, No phone services
glimpse(cleanData)
```

# Converting all variables to numeric form 
```{r}
# Converting all variables numeric form to be used in the models later which cannot use categorical variables
num.cleanData = data.frame(cleanData)

#num.cleanData$Churn  = as.numeric(num.cleanData$Churn)
num.cleanData$gender  = as.numeric(num.cleanData$gender)
num.cleanData$SeniorCitizen  = as.numeric(num.cleanData$SeniorCitizen)
num.cleanData$Partner = as.numeric(num.cleanData$Partner)
num.cleanData$Dependents  = as.numeric(num.cleanData$Dependents)
num.cleanData$tenure  = as.numeric(num.cleanData$tenure)
num.cleanData$PhoneService  = as.numeric(num.cleanData$PhoneService)
num.cleanData$Contract  = as.factor(num.cleanData$Contract)
num.cleanData$Contract  = as.numeric(num.cleanData$Contract)
num.cleanData$PaperlessBilling  = as.numeric(num.cleanData$PaperlessBilling)
num.cleanData$PaymentMethod  = as.factor(num.cleanData$PaymentMethod)
num.cleanData$PaymentMethod  = as.numeric(num.cleanData$PaymentMethod)
num.cleanData$MonthlyCharges  = as.numeric(num.cleanData$MonthlyCharges)
num.cleanData$TotalCharges  = as.numeric(num.cleanData$TotalCharges)
num.cleanData$MultipleLines  = as.numeric(num.cleanData$MultipleLines)
num.cleanData$InternetService  = as.factor(num.cleanData$InternetService)
num.cleanData$InternetService  = as.numeric(num.cleanData$InternetService)
num.cleanData$OnlineSecurity  = as.numeric(num.cleanData$OnlineSecurity)
num.cleanData$OnlineBackup  = as.numeric(num.cleanData$OnlineBackup)
num.cleanData$DeviceProtection  = as.numeric(num.cleanData$DeviceProtection)
num.cleanData$TechSupport  = as.numeric(num.cleanData$TechSupport)
num.cleanData$StreamingTV  = as.numeric(num.cleanData$StreamingTV)
num.cleanData$StreamingMovies  = as.numeric(num.cleanData$StreamingMovies)
glimpse(num.cleanData)

summary(num.cleanData)
```
```{r}
sapply(num.cleanData,class)
```




```{r}
#Don't execute below line of code twice
cleanData = cleanData[,c(12,1,2,3,4,5:20)] #bringing churn column at the 2nd place so that it is used in the correlation table below as reference
cleanData$Churn.1 = NULL #removing column Churn.1 which is a copy of churn column created after executing the above line of code

```

```{r}
#install.packages("DataExplorer")
library(DataExplorer)

plot_correlation(cleanData, type="c")

```


```{r}
sapply(cleanData,class)
```

```{r}
#creating testing and training datasets
set.seed(10)
data1 = sort(sample(nrow(cleanData), nrow(cleanData)*.70)) 
train.set1 = cleanData[data1,] 
test.set1 = cleanData[-data1,]
#data2 = sort(sample(nrow(cleanData), nrow(cleanData)*.60)) 
#train.set2 = cleanData[data2,] 
#test.set2 = cleanData[-data2,]

```


## logistic regression
```{r}
#install.packages("MASS")
library(MASS)
set.seed(10)
logistic.reg.model = glm(Churn ~., data = train.set1 , family="binomial")
summary(logistic.reg.model)
```


```{r}

pred1 = predict(logistic.reg.model, test.set1, type = "response")
pred1.glm = rep(0, length(pred1))
pred1.glm[pred1 > 0.5] = 1
table(pred1.glm, test.set1$Churn)
meanval1 = mean(pred1.glm != test.set1$Churn)
meanval1

cat("Test error of the Logistic Regression model is", meanval1*100,"%")
```


## Logistic with top 7 variables

```{r}
#install.packages("MASS")
library(MASS)
set.seed(10)
logistic.reg.model = glm(Churn ~ Contract+tenure+TotalCharges+MonthlyCharges+InternetService+PaymentMethod+PaperlessBilling , data = train.set1 , family="binomial")
summary(logistic.reg.model)
```


```{r}
#draw roc curve to find value of 0.5 (probablity threshold) or (probability, sensitivity, accuracy and specificity curver)
pred1 = predict(logistic.reg.model, test.set1, type = "response")
pred1.glm = rep(0, length(pred1))
pred1.glm[pred1 > 0.5] = 1
table(pred1.glm, test.set1$Churn)
meanval1 = mean(pred1.glm != test.set1$Churn)
meanval1

cat("Test error of the Logistic Regression model is", meanval1*100,"%")
```


## Lasso

```{r}
#install.packages("glmnet")
library(glmnet)
train.set1$Churn=as.numeric(train.set1$Churn)
test.set1$Churn=as.numeric(test.set1$Churn)
matrix.train.set1<- model.matrix(Churn ~ ., data = train.set1) 
matrix.test.set1 <- model.matrix(Churn ~ ., data = test.set1) 
grid <- 10 ^ seq(10, -2, length = 100) 
fit.lasso <- glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
set.seed(30)
cv.lasso <- cv.glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
bestlam.lasso <- cv.lasso$lambda.min
bestlam.lasso

```
```{r}
pred.lasso <- predict(fit.lasso, s = bestlam.lasso, newx = matrix.test.set1)
mean((pred.lasso - test.set1$Churn)^2)

```

```{r}
#finding non-zero coefficient estimates
predict(fit.lasso, s = bestlam.lasso, type = "coefficients")
```
There are 17 non-zero coefficient estimates as seen above after not considering Intercept and id


## Lasso with top 7 variables

```{r}
#install.packages("glmnet")
library(glmnet)
train.set1$Churn=as.numeric(train.set1$Churn)
test.set1$Churn=as.numeric(test.set1$Churn)
matrix.train.set1<- model.matrix(Churn ~ Contract+tenure+TotalCharges+MonthlyCharges+InternetService+PaymentMethod+PaperlessBilling , data = train.set1) 
matrix.test.set1 <- model.matrix(Churn ~ Contract+tenure+TotalCharges+MonthlyCharges+InternetService+PaymentMethod+PaperlessBilling , data = test.set1) 
grid <- 10 ^ seq(10, -2, length = 100) 
fit.lasso <- glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
set.seed(30)
cv.lasso <- cv.glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
bestlam.lasso <- cv.lasso$lambda.min
bestlam.lasso

```
```{r}
pred.lasso <- predict(fit.lasso, s = bestlam.lasso, newx = matrix.test.set1)
mean((pred.lasso - test.set1$Churn)^2)

```

```{r}
#finding non-zero coefficient estimates
predict(fit.lasso, s = bestlam.lasso, type = "coefficients")
```
Accuracy with top 6 var is 85.96



## KNN

```{r}
set.seed(10)
num.data1 = sort(sample(nrow(num.cleanData), nrow(num.cleanData)*.70)) 
num.train.set1 = num.cleanData[num.data1,] 
num.test.set1 = num.cleanData[-num.data1,]


KNNtrain1 = cbind(num.train.set1$gender, num.train.set1$SeniorCitizen, num.train.set1$Partner, num.train.set1$Dependents, num.train.set1$tenure, num.train.set1$PhoneService, num.train.set1$Contract,num.train.set1$PaperlessBilling ,num.train.set1$PaymentMethod, num.train.set1$MonthlyCharges, num.train.set1$TotalCharges, num.train.set1$MultipleLines, num.train.set1$InternetService, num.train.set1$OnlineSecurity, num.train.set1$OnlineBackup, num.train.set1$DeviceProtection, num.train.set1$TechSupport, num.train.set1$StreamingTV, num.train.set1$StreamingMovies)

KNNtest1 = cbind(num.test.set1$gender, num.test.set1$SeniorCitizen, num.test.set1$Partner, num.test.set1$Dependents, num.test.set1$tenure, num.test.set1$PhoneService, num.test.set1$Contract ,num.test.set1$PaperlessBilling ,num.test.set1$PaymentMethod, num.test.set1$MonthlyCharges, num.test.set1$TotalCharges, num.test.set1$MultipleLines, num.test.set1$InternetService, num.test.set1$OnlineSecurity, num.test.set1$OnlineBackup, num.test.set1$DeviceProtection, num.test.set1$TechSupport, num.test.set1$StreamingTV, num.test.set1$StreamingMovies)

KNN_Col1 = cbind("gender","SeniorCitizen","Partner", "Dependents", "tenure", "PhoneService", "Contract", "PaperlessBilling", "PaymentMethod", "MonthlyCharges", "TotalCharges", "MultipleLines", "InternetService","OnlineSecurity","OnlineBackup","DeviceProtection","TechSupport","StreamingTV","StreamingMovies")
colnames(KNNtrain1) = KNN_Col1
colnames(KNNtest1) = KNN_Col1
```
```{r}
library(class)
num.train.set1.Churn = num.train.set1$Churn
set.seed(1)

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 5)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K= 5")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 10)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=10")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 15)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=15")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 20)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=20")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 25)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=25")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 30)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=30")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 35)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=35")
```

## KNN with top 7 variables

```{r}
set.seed(10)
num.data1 = sort(sample(nrow(num.cleanData), nrow(num.cleanData)*.70)) 
num.train.set1 = num.cleanData[num.data1,] 
num.test.set1 = num.cleanData[-num.data1,]


KNNtrain1 = cbind(num.train.set1$tenure, num.train.set1$Contract, num.train.set1$PaymentMethod, num.train.set1$MonthlyCharges, num.train.set1$TotalCharges,  num.train.set1$InternetService, num.train.set1$PaperlessBilling )

KNNtest1 = cbind(num.test.set1$tenure, num.test.set1$Contract, num.test.set1$PaymentMethod, num.test.set1$MonthlyCharges, num.test.set1$TotalCharges,  num.test.set1$InternetService, num.test.set1$PaperlessBilling)

KNN_Col1 = cbind("tenure", "Contract", "PaymentMethod", "MonthlyCharges", "TotalCharges",  "InternetService","PaperlessBilling")
colnames(KNNtrain1) = KNN_Col1
colnames(KNNtest1) = KNN_Col1
```


```{r}
library(class)
num.train.set1.Churn = num.train.set1$Churn
set.seed(1)

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 5)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K= 5")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 10)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=10")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 15)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=15")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 20)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=20")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 25)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=25")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 30)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=30")

kNNModel1 = knn(KNNtrain1, KNNtest1, num.train.set1.Churn, k = 35)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != num.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=35")
```



## Random Forest
```{r}
#applying random forest
library(randomForest)
cleanData$Churn = as.factor(cleanData$Churn)
data1 = sort(sample(nrow(cleanData), nrow(cleanData)*.70)) 
train.set1 = cleanData[data1,] 
test.set1 = cleanData[-data1,]

#levels(test.set1$customerID)<-levels(train.set1$customerID)
levels(test.set1$Churn)<-levels(train.set1$Churn)
levels(test.set1$gender)<-levels(train.set1$gender)
levels(test.set1$SeniorCitizen)<-levels(train.set1$SeniorCitizen)
levels(test.set1$Partner)<-levels(train.set1$Partner)
levels(test.set1$Dependents)<-levels(train.set1$Dependents)
levels(test.set1$tenure)<-levels(train.set1$tenure)
levels(test.set1$PhoneService)<-levels(train.set1$PhoneService)
levels(test.set1$Contract)<-levels(train.set1$Contract)
levels(test.set1$PaperlessBilling)<-levels(train.set1$PaperlessBilling)
levels(test.set1$PaymentMethod)<-levels(train.set1$PaymentMethod)
levels(test.set1$MonthlyCharges)<-levels(train.set1$MonthlyCharges)
levels(test.set1$TotalCharges)<-levels(train.set1$TotalCharges)
levels(test.set1$MultipleLines)<-levels(train.set1$MultipleLines)
levels(test.set1$InternetService)<-levels(train.set1$InternetService)
levels(test.set1$OnlineSecurity)<-levels(train.set1$OnlineSecurity)
levels(test.set1$OnlineBackup)<-levels(train.set1$OnlineBackup)
levels(test.set1$DeviceProtection)<-levels(train.set1$DeviceProtection)
levels(test.set1$TechSupport)<-levels(train.set1$TechSupport)
levels(test.set1$StreamingTV)<-levels(train.set1$StreamingTV)
levels(test.set1$StreamingMovies)<-levels(train.set1$StreamingMovies)
levels(test.set1$Contract)<-levels(train.set1$Contract)



set.seed(123)
y_test <- cleanData["Churn"]
fit1 <- randomForest(Churn~. , data=train.set1, mtry = 19, nTree=500, importance=TRUE)
fit1
y_pred=predict(fit1,newdata = test.set1)

#Using IMportance to find out which variables are important
importance(fit1)

t1=table(y_pred, test.set1$Churn)
Accuracy = (t1[1,1] + t1[2,2]) / (dim(test.set1)[1])
cat("Accuracy is ", Accuracy)
```



## Random forest with top 7 variables
```{r}
#applying random forest
library(randomForest)

t7.cleanData = cleanData[c("Churn", "Contract","tenure","TotalCharges","MonthlyCharges","InternetService","PaymentMethod","PaperlessBilling")]
set.seed(1)


t7.data1 = sort(sample(nrow(t7.cleanData), nrow(t7.cleanData)*.70)) 
t7.train.set1 = t7.cleanData[t7.data1,] 
t7.test.set1 = t7.cleanData[-t7.data1,]

#levels(test.set1$customerID)<-levels(train.set1$customerID)
levels(t7.test.set1$Churn)<-levels(t7.train.set1$Churn)
levels(t7.test.set1$tenure)<-levels(t7.train.set1$tenure)
levels(t7.test.set1$Contract)<-levels(t7.train.set1$Contract)
levels(t7.test.set1$PaymentMethod)<-levels(t7.train.set1$PaymentMethod)
levels(t7.test.set1$MonthlyCharges)<-levels(t7.train.set1$MonthlyCharges)
levels(t7.test.set1$TotalCharges)<-levels(t7.train.set1$TotalCharges)
levels(t7.test.set1$InternetService)<-levels(t7.train.set1$InternetService) 
levels(t7.test.set1$PaperlessBilling)<-levels(t7.train.set1$PaperlessBilling)


set.seed(123)
fit1 <- randomForest(Churn~. , data=t7.train.set1, mtry = 7, nTree=500, importance=TRUE)
fit1
y_pred=predict(fit1,newdata = t7.test.set1)

#Using IMportance to find out which variables are important
importance(fit1)

t1=table(y_pred, t7.test.set1$Churn)
Accuracy = (t1[1,1] + t1[2,2]) / (dim(t7.test.set1)[1])
cat("Accuracy is ", Accuracy)
```

## Bagging Model
```{r}
#install.packages("randomForest")
library(randomForest)

#I was getting this error "Type of predictors in new data do not match types of the training data" from predict() line of code. As a solution, I found that by adding the below line of codes until set.seed(1), it resolves the issues. Source : https://stackoverflow.com/questions/24829674/r-random-forest-error-type-of-predictors-in-new-data-do-not-match

levels(test.set1$Churn)<-levels(train.set1$Churn)
levels(test.set1$gender)<-levels(train.set1$gender)
levels(test.set1$SeniorCitizen)<-levels(train.set1$SeniorCitizen)
levels(test.set1$Partner)<-levels(train.set1$Partner)
levels(test.set1$Dependents)<-levels(train.set1$Dependents)
levels(test.set1$tenure)<-levels(train.set1$tenure)
levels(test.set1$PhoneService)<-levels(train.set1$PhoneService)
levels(test.set1$Contract)<-levels(train.set1$Contract)
levels(test.set1$PaperlessBilling)<-levels(train.set1$PaperlessBilling)
levels(test.set1$PaymentMethod)<-levels(train.set1$PaymentMethod)
levels(test.set1$MonthlyCharges)<-levels(train.set1$MonthlyCharges)
levels(test.set1$TotalCharges)<-levels(train.set1$TotalCharges)
levels(test.set1$MultipleLines)<-levels(train.set1$MultipleLines)
levels(test.set1$InternetService)<-levels(train.set1$InternetService)
levels(test.set1$OnlineSecurity)<-levels(train.set1$OnlineSecurity)
levels(test.set1$OnlineBackup)<-levels(train.set1$OnlineBackup)
levels(test.set1$DeviceProtection)<-levels(train.set1$DeviceProtection)
levels(test.set1$TechSupport)<-levels(train.set1$TechSupport)
levels(test.set1$StreamingTV)<-levels(train.set1$StreamingTV)
levels(test.set1$StreamingMovies)<-levels(train.set1$StreamingMovies)
levels(test.set1$Contract)<-levels(train.set1$Contract)


test.set1 = rbind(train.set1[1, ] , test.set1)
test.set1 = test.set1[-1,]

set.seed(1)
fit.bagging = randomForest(Churn~., data=train.set1, mtry = 19, importance=TRUE)
pred2=predict(fit.bagging,newdata = test.set1)

bagging.table=table(pred2, test.set1$Churn)
accuracy = (bagging.table[1,1] + bagging.table[2,2]) / (dim(test.set1)[1])
cat("Accuracy of the Bagging Model is", accuracy*100,"%")
```

```{r}
#Using importance function to find important variables
importance(fit.bagging)
varImpPlot(fit.bagging)
```


## Bagging Model with top 7 variables
```{r}
#install.packages("randomForest")
library(randomForest)

#I was getting this error "Type of predictors in new data do not match types of the training data" from predict() line of code. As a solution, I found that by adding the below line of codes until set.seed(1), it resolves the issues. Source : https://stackoverflow.com/questions/24829674/r-random-forest-error-type-of-predictors-in-new-data-do-not-match

#levels(test.set1$customerID)<-levels(train.set1$customerID)
levels(t7.test.set1$Churn)<-levels(t7.train.set1$Churn)
levels(t7.test.set1$tenure)<-levels(t7.train.set1$tenure)
levels(t7.test.set1$Contract)<-levels(t7.train.set1$Contract)
levels(t7.test.set1$PaymentMethod)<-levels(t7.train.set1$PaymentMethod)
levels(t7.test.set1$MonthlyCharges)<-levels(t7.train.set1$MonthlyCharges)
levels(t7.test.set1$TotalCharges)<-levels(t7.train.set1$TotalCharges)
levels(t7.test.set1$InternetService)<-levels(t7.train.set1$InternetService) 
levels(t7.test.set1$PaperlessBilling)<-levels(t7.train.set1$PaperlessBilling)


t7.test.set1 = rbind(t7.train.set1[1, ] , t7.test.set1)
t7.test.set1 = t7.test.set1[-1,]

set.seed(1)
fit.bagging = randomForest(Churn~., data=t7.train.set1, mtry = 7, importance=TRUE)
pred2=predict(fit.bagging,newdata = t7.test.set1)

bagging.table=table(pred2, t7.test.set1$Churn)
accuracy = (bagging.table[1,1] + bagging.table[2,2]) / (dim(t7.test.set1)[1])
cat("Accuracy of the Bagging Model is", accuracy*100,"%")
```

```{r}
#Using importance function to find important variables
importance(fit.bagging)
varImpPlot(fit.bagging)
```


## Boosting Model
```{r}
#install.packages("gbm")
library(gbm)
set.seed(1)
#converting response variable to char type to be used in the boosting model
train.set1$Churn = as.character(train.set1$Churn)
test.set1$Churn = as.character(test.set1$Churn)
fit.boost=gbm(Churn∼., data = train.set1, distribution = "bernoulli", n.trees=500, interaction.depth=4, shrinkage = 0.1)
pred4 = predict(fit.boost, newdata = test.set1, n.trees = 500)
pred4 = ifelse(pred4 > 0.5,1,0)
boost.table = table(pred4, test.set1$Churn)

accuracy = (boost.table[1,1] + boost.table[2,2]) / (dim(test.set1)[1])
cat("Accuracy of the Boosting Model is", accuracy*100)
```  

```{r}
#Using summary function to find important variables
summary(fit.boost)
```


## Boosting Model with top 7 variables
```{r}
#install.packages("gbm")
library(gbm)
set.seed(1)
#converting response variable to char type to be used in the boosting model
t7.train.set1$Churn = as.character(t7.train.set1$Churn)
t7.test.set1$Churn = as.character(t7.test.set1$Churn)
fit.boost=gbm(Churn∼., data = t7.train.set1, distribution = "bernoulli", n.trees=500, interaction.depth=4, shrinkage = 0.1)
pred4 = predict(fit.boost, newdata = t7.test.set1, n.trees = 500)
pred4 = ifelse(pred4 > 0.5,1,0)
boost.table = table(pred4, t7.test.set1$Churn)

accuracy = (boost.table[1,1] + boost.table[2,2]) / (dim(t7.test.set1)[1])
cat("Accuracy of the Boosting Model is", accuracy*100)
```  

```{r}
#Using summary function to find important variables
summary(fit.boost)
```


## SVM with linear kernel

```{r}
library(e1071)
num.train.set1$Churn = as.factor(num.train.set1$Churn)
num.test.set1$Churn = as.factor(num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = num.train.set1, kernel = "linear", cost = 0.01)
summary(svc.fit)
```

the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,num.train.set1)
table(num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate is",(633+358)/(3254+685+633+358))
```

```{r}
test.pred <- predict(svc.fit, num.test.set1)
table(num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate is",(268+137)/(1425+283+268+137))
```

## SVM with polynomial kernel degree=2

```{r}
library(e1071)
num.train.set1$Churn = as.factor(num.train.set1$Churn)
num.test.set1$Churn = as.factor(num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = num.train.set1, kernel = "polynomial", degree=2)
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,num.train.set1)
table(num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate is",(700+270)/(3342+618+700+270))
```

```{r}
test.pred <- predict(svc.fit, num.test.set1)
table(num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate is",(309+110)/(1452+242+309+110))
```

## SVM with polynomial kernel degree=3

```{r}
library(e1071)
num.train.set1$Churn = as.factor(num.train.set1$Churn)
num.test.set1$Churn = as.factor(num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = num.train.set1, kernel = "polynomial", degree=3)
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,num.train.set1)
table(num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate is",(600+254)/(3358+718+600+254))
```

```{r}
test.pred <- predict(svc.fit, num.test.set1)
table(num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate is",(146+290)/(1416+261+146+290))
```

## SVM with radial kernel

```{r}
library(e1071)
num.train.set1$Churn = as.factor(num.train.set1$Churn)
num.test.set1$Churn = as.factor(num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = num.train.set1, kernel = "radial")
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,num.train.set1)
table(num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate is",(617+266)/(3346+701+617+266))
```

```{r}
test.pred <- predict(svc.fit, num.test.set1)
table(num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate is",(291+127)/(1435+260+291+127))
```


## SVM with top 7 variables

```{r}
t7.num.cleanData = num.cleanData[c("Churn","Contract","tenure","TotalCharges","MonthlyCharges","InternetService","PaymentMethod","PaperlessBilling")]
set.seed(1)


t7.num.data1 = sort(sample(nrow(t7.num.cleanData), nrow(t7.num.cleanData)*.70)) 
t7.num.train.set1 = t7.num.cleanData[t7.num.data1,] 
t7.num.test.set1 = t7.num.cleanData[-t7.num.data1,]

```


## SVM with linear kernel top 7 variables

```{r}
library(e1071)
t7.num.train.set1$Churn = as.factor(t7.num.train.set1$Churn)
t7.num.test.set1$Churn = as.factor(t7.num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = t7.num.train.set1, kernel = "linear", cost = 0.01)
summary(svc.fit)
```

the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,t7.num.train.set1)
table(t7.num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate with top 7 variable is",(694+330)/(3310+596+694+330))
```

```{r}
test.pred <- predict(svc.fit, t7.num.test.set1)
table(t7.num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate with top 7 variable is",(305+151)/(1383+274+305+151))
```

## SVM with polynomial kernel degree=2  with  7 variables

```{r}
library(e1071)
t7.num.train.set1$Churn = as.factor(t7.num.train.set1$Churn)
t7.num.test.set1$Churn = as.factor(t7.num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = t7.num.train.set1, kernel = "polynomial", degree=2)
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,t7.num.train.set1)
table(t7.num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate with top 7 variable is",(861+182)/(3458+429+861+182))
```

```{r}
test.pred <- predict(svc.fit, t7.num.test.set1)
table(t7.num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate with top 7 variable is",(382+83)/(1451+199+380+83))
```

## SVM with polynomial kernel degree=3  with top 7 variables

```{r}
library(e1071)
t7.num.train.set1$Churn = as.factor(t7.num.train.set1$Churn)
t7.num.test.set1$Churn = as.factor(t7.num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = t7.num.train.set1, kernel = "polynomial", degree=3)
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,t7.num.train.set1)
table(t7.num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate with top 7 variable is",(793+215)/(3425+497+793+215))
```

```{r}
test.pred <- predict(svc.fit, t7.num.test.set1)
table(t7.num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate with top 7 variable is",(345+103)/(1431+234+345+103))
```

## SVM with radial kernel  with top 7 variables

```{r}
library(e1071)
t7.num.train.set1$Churn = as.factor(t7.num.train.set1$Churn)
t7.num.test.set1$Churn = as.factor(t7.num.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = t7.num.train.set1, kernel = "radial")
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,t7.num.train.set1)
table(t7.num.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate with top 6 variable is",(665+344)/(3296+625+665+334))
```

```{r}
test.pred <- predict(svc.fit, t7.num.test.set1)
table(t7.num.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate with top 6 variable is",(300+150)/(1384+279+300+150))
```




# creating dummy variables for the variables who are in factor form (one-hot encoding)
```{r}
#creating a new dataset d.cleanData where we will add dummy variables for the categorical variables
d.cleanData = data.frame(cleanData)

for(unique_value in unique(d.cleanData$Contract)){
d.cleanData[paste("Contract", unique_value, sep = ".")] <- ifelse(d.cleanData$Contract == unique_value, 1, 0)
}

for(unique_value in unique(d.cleanData$PaymentMethod)){
d.cleanData[paste("PaymentMethod", unique_value, sep = ".")] <- ifelse(d.cleanData$PaymentMethod == unique_value, 1, 0)
}

for(unique_value in unique(d.cleanData$InternetService)){
d.cleanData[paste("InternetService", unique_value, sep = ".")] <- ifelse(d.cleanData$InternetService == unique_value, 1, 0)
}
#removing Contract, PaymentMethod, InternetService from the dataset as we have created dummy variables for those 3 variables above
d.cleanData$Contract=NULL

d.cleanData$PaymentMethod=NULL

d.cleanData$InternetService=NULL

```


```{r}
colnames(d.cleanData)
```


```{r}
d.cleanData$Churn = as.factor(d.cleanData$Churn)
sapply(d.cleanData,class)
```

```{r}

#install.packages("DataExplorer")
library(DataExplorer)

plot_correlation(d.cleanData, type="c")

```


```{r}

d.cleanData$Churn = as.factor(d.cleanData$Churn)
d.data1 = sort(sample(nrow(d.cleanData), nrow(d.cleanData)*.70)) 
d.train.set1 = d.cleanData[d.data1,] 
d.test.set1 = d.cleanData[-d.data1,]

```


```{r}
colnames(d.train.set1)
```




## Logistic with top 7 variables (using dummy variables)

```{r}
#install.packages("MASS")
library(MASS)
set.seed(10)
d.t7.logistic.reg.model = glm(Churn ~  tenure  + MonthlyCharges  + TotalCharges  +  Contract.One.year  +  Contract.Monthtomonth  +  Contract.Two.year  +  PaperlessBilling  + PaymentMethod.Mailed.check  + PaymentMethod.Electronic.check  + PaymentMethod.Credit.card.automatic  + PaymentMethod.Bank.transfer.automatic  + InternetService.DSL  +  InternetService.Fiber.optic  +  InternetService.No , data = d.train.set1 , family="binomial")
summary(d.t7.logistic.reg.model)
```


```{r}
d.t7.pred1 = predict(d.t7.logistic.reg.model, d.test.set1, type = "response")
d.t7.pred1.glm = rep(0, length(d.t7.pred1))
d.t7.pred1.glm[d.t7.pred1 > 0.5] = 1
table(d.t7.pred1.glm, d.test.set1$Churn)
d.t7.meanval1 = mean(d.t7.pred1.glm != d.test.set1$Churn)
d.t7.meanval1

cat("Test error of the Logistic Regression model is", d.t7.meanval1*100,"%")
```


## KNN with top 7 variables (using dummy variables)

```{r}

set.seed(10)
#d.data1 = sort(sample(nrow(d.cleanData), nrow(d.cleanData)*.70)) 
#d.train.set1 = d.cleanData[d.data1,] 
#d.test.set1 = d.cleanData[-d.data1,]

d.t7.KNNtrain1 = cbind(d.train.set1$tenure , d.train.set1$MonthlyCharges , d.train.set1$TotalCharges , d.train.set1$Contract.One.year , d.train.set1$Contract.Monthtomonth, d.train.set1$Contract.Two.year , d.train.set1$PaperlessBilling , d.train.set1$PaymentMethod.Mailed.check , d.train.set1$PaymentMethod.Electronic.check , d.train.set1$PaymentMethod.Credit.card.automatic , d.train.set1$PaymentMethod.Bank.transfer.automatic , d.train.set1$InternetService.DSL , d.train.set1$InternetService.Fiber.optic , d.train.set1$InternetService.No)

d.t7.KNNtest1 = cbind(d.test.set1$tenure , d.test.set1$MonthlyCharges , d.test.set1$TotalCharges , d.test.set1$Contract.One.year , d.test.set1$Contract.Monthtomonth, d.test.set1$Contract.Two.year , d.test.set1$PaperlessBilling , d.test.set1$PaymentMethod.Mailed.check , d.test.set1$PaymentMethod.Electronic.check , d.test.set1$PaymentMethod.Credit.card.automatic , d.test.set1$PaymentMethod.Bank.transfer.automatic , d.test.set1$InternetService.DSL , d.test.set1$InternetService.Fiber.optic , d.test.set1$InternetService.No)

d.t7.KNN_Col1 = cbind("tenure","MonthlyCharges","TotalCharges", "Contract.One.year", "Contract.Monthtomonth", "Contract.Two.year", "PaperlessBilling","PaymentMethod.Mailed.check","PaymentMethod.Electronic.check","PaymentMethod.Credit.card.automatic","PaymentMethod.Bank.transfer.automatic","InternetService.DSL", "InternetService.Fiber.optic", "InternetService.No")
colnames(d.t7.KNNtrain1) = d.t7.KNN_Col1
colnames(d.t7.KNNtest1) = d.t7.KNN_Col1
```


```{r}
library(class)
d.t7.train.set1.Churn = d.train.set1$Churn
set.seed(1)

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 5)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K= 5")

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 10)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=10")

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 15)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=15")

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 20)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=20")

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 25)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=25")

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 30)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=30")

kNNModel1 = knn(d.t7.KNNtrain1, d.t7.KNNtest1, d.t7.train.set1.Churn, k = 35)
table(kNNModel1, num.test.set1$Churn)
meanval1 = mean(kNNModel1 != d.test.set1$Churn)
meanval1
cat("Test error of the KNN model is", meanval1*100,"% for K=35")
```


## Lasso with top 7 variables (using dummy variables)

```{r}
#install.packages("glmnet")
library(glmnet)
#d.data1 = sort(sample(nrow(d.cleanData), nrow(d.cleanData)*.70)) 
#d.train.set1 = d.cleanData[d.data1,] 
#d.test.set1 = d.cleanData[-d.data1,]
d.train.set1$Churn=as.numeric(d.train.set1$Churn)
d.test.set1$Churn=as.numeric(d.test.set1$Churn)

d.t7.matrix.train.set1<- model.matrix(Churn ~ tenure  + MonthlyCharges  + TotalCharges  +  Contract.One.year  +  Contract.Monthtomonth  +  Contract.Two.year  +  PaperlessBilling  + PaymentMethod.Mailed.check  + PaymentMethod.Electronic.check  + PaymentMethod.Credit.card.automatic  + PaymentMethod.Bank.transfer.automatic  + InternetService.DSL  +  InternetService.Fiber.optic  +  InternetService.No , data = d.train.set1) 

d.t7.matrix.test.set1 <- model.matrix(Churn ~ tenure  + MonthlyCharges  + TotalCharges  +  Contract.One.year  +  Contract.Monthtomonth  +  Contract.Two.year  +  PaperlessBilling  + PaymentMethod.Mailed.check  + PaymentMethod.Electronic.check  + PaymentMethod.Credit.card.automatic  + PaymentMethod.Bank.transfer.automatic  + InternetService.DSL  +  InternetService.Fiber.optic  +  InternetService.No , data = d.test.set1) 

grid <- 10 ^ seq(10, -2, length = 100) 

fit.lasso <- glmnet(d.t7.matrix.train.set1, d.train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
set.seed(30)
cv.lasso <- cv.glmnet(d.t7.matrix.train.set1, d.train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
bestlam.lasso <- cv.lasso$lambda.min
bestlam.lasso

```
```{r}
pred.lasso <- predict(fit.lasso, s = bestlam.lasso, newx = d.t7.matrix.test.set1)
mean((pred.lasso - d.test.set1$Churn)^2)

```

```{r}
#finding non-zero coefficient estimates
predict(fit.lasso, s = bestlam.lasso, type = "coefficients")
```

As per LASSO, important variables are tenure, MonthlyCharges, TotalCharges, Contract.Monthtomonth, PaperlessBilling, PaymentMethod.Mailed.check, PaymentMethod.Electronic.check, InternetService.Fiber.optic, InternetService.No



## Random Forest Model with top 7 (using dummy variables)
```{r}
library(randomForest)
#d.data1 = sort(sample(nrow(d.cleanData), nrow(d.cleanData)*.70)) 
#d.train.set1 = d.cleanData[d.data1,] 
#d.test.set1 = d.cleanData[-d.data1,]

d.t7.cleanData = d.cleanData[c("Churn","tenure","MonthlyCharges","TotalCharges", "Contract.One.year", "Contract.Monthtomonth", "Contract.Two.year", "PaperlessBilling","PaymentMethod.Mailed.check","PaymentMethod.Electronic.check","PaymentMethod.Credit.card.automatic","PaymentMethod.Bank.transfer.automatic","InternetService.DSL", "InternetService.Fiber.optic", "InternetService.No")]
set.seed(1)


d.t7.data1 = sort(sample(nrow(d.t7.cleanData), nrow(d.t7.cleanData)*.70)) 
d.t7.train.set1 = d.t7.cleanData[d.t7.data1,] 
d.t7.test.set1 = d.t7.cleanData[-d.t7.data1,]

#levels(test.set1$customerID)<-levels(train.set1$customerID)
levels(d.t7.test.set1$Churn)<-levels(d.t7.train.set1$Churn)
levels(d.t7.test.set1$tenure)<-levels(d.t7.train.set1$tenure)
levels(d.t7.test.set1$PaperlessBilling)<-levels(d.t7.train.set1$PaperlessBilling)
levels(d.t7.test.set1$PaymentMethod.Mailed.check)<-levels(d.t7.train.set1$PaymentMethod.Mailed.check)
levels(d.t7.test.set1$PaymentMethod.Electronic.check)<-levels(d.t7.train.set1$PaymentMethod.Electronic.check)
levels(d.t7.test.set1$PaymentMethod.Credit.card.automatic)<-levels(d.t7.train.set1$PaymentMethod.Credit.card.automatic)
levels(d.t7.test.set1$PaymentMethod.Bank.transfer.automatic)<-levels(d.t7.train.set1$PaymentMethod.Bank.transfer.automatic)
levels(d.t7.test.set1$MonthlyCharges)<-levels(d.t7.train.set1$MonthlyCharges)
levels(d.t7.test.set1$TotalCharges)<-levels(d.t7.train.set1$TotalCharges)
levels(d.t7.test.set1$InternetService.DSL)<-levels(d.t7.train.set1$InternetService.DSL)
levels(d.t7.test.set1$InternetService.Fiber.optic)<-levels(d.t7.train.set1$InternetService.Fiber.optic)
levels(d.t7.test.set1$InternetService.No)<-levels(d.t7.train.set1$InternetService.No)
levels(d.t7.test.set1$Contract.One.year)<-levels(d.t7.train.set1$Contract.One.year)
levels(d.t7.test.set1$Contract.Monthtomonth)<-levels(d.t7.train.set1$Contract.Monthtomonth)
levels(d.t7.test.set1$Contract.Two.year)<-levels(d.t7.train.set1$Contract.Two.year)


set.seed(123)
y_test <- d.t7.test.set1["Churn"]
fit1 <- randomForest(Churn~ tenure  + MonthlyCharges  + TotalCharges  +  Contract.One.year  +  Contract.Monthtomonth  +  Contract.Two.year  +  PaperlessBilling  + PaymentMethod.Mailed.check  + PaymentMethod.Electronic.check  + PaymentMethod.Credit.card.automatic  + PaymentMethod.Bank.transfer.automatic  + InternetService.DSL  +  InternetService.Fiber.optic  +  InternetService.No , data=d.t7.train.set1, mtry = 14, nTree=500, importance=TRUE)
fit1
y_pred=predict(fit1,newdata = d.t7.test.set1)

#Using IMportance to find out which variables are important

importance(fit1)
varImpPlot(fit1)
t1=table(y_pred, d.t7.test.set1$Churn)
Accuracy = (t1[1,1] + t1[2,2]) / (dim(d.t7.test.set1)[1])
cat("Accuracy is ", Accuracy)
```



## Boosting Model with top 7 (using dummy variables)
```{r}
#install.packages("gbm")
library(gbm)
set.seed(1)
#converting response variable to char type to be used in the boosting model
d.t7.train.set1$Churn = as.character(d.t7.train.set1$Churn)
d.t7.test.set1$Churn = as.character(d.t7.test.set1$Churn)
fit.boost=gbm(Churn∼., data = d.t7.train.set1, distribution = "bernoulli", n.trees=500, interaction.depth=4, shrinkage = 0.1)
pred4 = predict(fit.boost, newdata = d.t7.test.set1, n.trees = 500)
pred4 = ifelse(pred4 > 0.5,1,0)
boost.table = table(pred4, d.t7.test.set1$Churn)

accuracy = (boost.table[1,1] + boost.table[2,2]) / (dim(d.t7.test.set1)[1])
cat("Accuracy of the Boosting Model is", accuracy*100)
```

```{r}
#Using summary function to find important variables
summary(fit.boost)
```




## SVM with top 7 (using dummy variables)

```{r}

d.t7.cleanData = d.cleanData[c("Churn","tenure","MonthlyCharges","TotalCharges", "Contract.One.year", "Contract.Monthtomonth", "Contract.Two.year", "PaperlessBilling","PaymentMethod.Mailed.check","PaymentMethod.Electronic.check","PaymentMethod.Credit.card.automatic","PaymentMethod.Bank.transfer.automatic","InternetService.DSL", "InternetService.Fiber.optic", "InternetService.No")]
set.seed(1)


d.t7.data1 = sort(sample(nrow(d.t7.cleanData), nrow(d.t7.cleanData)*.70)) 
d.t7.train.set1 = d.t7.cleanData[d.t7.data1,] 
d.t7.test.set1 = d.t7.cleanData[-d.t7.data1,]

```



## SVM with linear kernel  top 7 (using dummy variables)

```{r}
library(e1071)
d.t7.train.set1$Churn = as.factor(d.t7.train.set1$Churn)
d.t7.test.set1$Churn = as.factor(d.t7.test.set1$Churn)
set.seed(123)
svc.fit <- svm(Churn ~ ., data = d.t7.train.set1, kernel = "linear")
summary(svc.fit)
```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,d.t7.train.set1)
table(d.t7.train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate with top 7 variable is",(632+390)/(3250+658+632+390))
```

```{r}
test.pred <- predict(svc.fit, d.t7.test.set1)
table(d.t7.test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate with top 7 variable is",(281+175)/(1359+298+281+175))
```



# SMOTE Sampling


```{r}
as.data.frame(table(d.cleanData$Churn))
```
```{r}
library(DMwR)
set.seed(1)
## Smote : Synthetic Minority Oversampling Technique To Handle Class Imbalancy In Binary Classification
newdata <- SMOTE(Churn ~., cleanData, perc.over = 100)

as.data.frame(table(newdata$Churn))
```

## Boosting model with SMOTE sampling
```{r}
library(gbm)
set.seed(1)

data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
train.set1 = newdata[data1,] 
test.set1 = newdata[-data1,]
#converting response variable to char type to be used in the boosting model
train.set1$Churn = as.character(train.set1$Churn)
test.set1$Churn = as.character(test.set1$Churn)
fit.boost=gbm(Churn∼., data = train.set1, distribution = "bernoulli", n.trees=500, interaction.depth=4, shrinkage = 0.1)
pred4 = predict(fit.boost, newdata = test.set1, n.trees = 500)
pred4 = ifelse(pred4 > 0.5,1,0)
boost.table = table(pred4, test.set1$Churn)

accuracy = (boost.table[1,1] + boost.table[2,2]) / (dim(test.set1)[1])
cat("Accuracy of the Boosting Model is", accuracy*100)
```

## Random forest model with SMOTE sampling
```{r}
library(randomForest)
set.seed(1)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
d.train.set1 = newdata[d.data1,] 
d.test.set1 = newdata[-d.data1,]


#levels(test.set1$customerID)<-levels(train.set1$customerID)
levels(d.test.set1$Churn)<-levels(d.train.set1$Churn)
levels(d.test.set1$gender)<-levels(d.train.set1$gender)
levels(d.test.set1$SeniorCitizen)<-levels(d.train.set1$SeniorCitizen)
levels(d.test.set1$Partner)<-levels(d.train.set1$Partner)
levels(d.test.set1$Dependents)<-levels(d.train.set1$Dependents)
levels(d.test.set1$tenure)<-levels(d.train.set1$tenure)
levels(d.test.set1$PhoneService)<-levels(d.train.set1$PhoneService)
#levels(test.set1$Contract)<-levels(train.set1$Contract)
levels(d.test.set1$PaperlessBilling)<-levels(d.train.set1$PaperlessBilling)
levels(d.test.set1$PaymentMethod.Mailed.check)<-levels(d.train.set1$PaymentMethod.Mailed.check)
levels(d.test.set1$PaymentMethod.Electronic.check)<-levels(d.train.set1$PaymentMethod.Electronic.check)
levels(d.test.set1$PaymentMethod.Credit.card.automatic)<-levels(d.train.set1$PaymentMethod.Credit.card.automatic)
levels(d.test.set1$PaymentMethod.Bank.transfer.automatic)<-levels(d.train.set1$PaymentMethod.Bank.transfer.automatic)
levels(d.test.set1$MonthlyCharges)<-levels(d.train.set1$MonthlyCharges)
levels(d.test.set1$TotalCharges)<-levels(d.train.set1$TotalCharges)
levels(d.test.set1$MultipleLines)<-levels(d.train.set1$MultipleLines)
levels(d.test.set1$InternetService.DSL)<-levels(d.train.set1$InternetService.DSL)
levels(d.test.set1$InternetService.Fiber.optic)<-levels(d.train.set1$InternetService.Fiber.optic)
levels(d.test.set1$InternetService.No)<-levels(d.train.set1$InternetService.No)
levels(d.test.set1$OnlineSecurity)<-levels(d.train.set1$OnlineSecurity)
levels(d.test.set1$OnlineBackup)<-levels(d.train.set1$OnlineBackup)
levels(d.test.set1$DeviceProtection)<-levels(d.train.set1$DeviceProtection)
levels(d.test.set1$TechSupport)<-levels(d.train.set1$TechSupport)
levels(d.test.set1$StreamingTV)<-levels(d.train.set1$StreamingTV)
levels(d.test.set1$StreamingMovies)<-levels(d.train.set1$StreamingMovies)
levels(d.test.set1$Contract.One.year)<-levels(d.train.set1$Contract.One.year)
levels(d.test.set1$Contract.Monthtomonth)<-levels(d.train.set1$Contract.Monthtomonth)
levels(d.test.set1$Contract.Two.year)<-levels(d.train.set1$Contract.Two.year)

y_test <- d.test.set1["Churn"]
fit1 <- randomForest(Churn ~. , data=d.train.set1, mtry = 26, nTree=500, importance=TRUE)
fit1
y_pred=predict(fit1,newdata = d.test.set1)

#Using IMportance to find out which variables are important
importance(fit1)

t1=table(y_pred, d.test.set1$Churn)
Accuracy = (t1[1,1] + t1[2,2]) / (dim(d.test.set1)[1])
cat("Accuracy is ", Accuracy)
```

```{r}
library(ROCR)
library(pROC)
# need to create prediction object from ROCR
pr <- prediction(pred4, test.set1$Churn)
#rf.roc <- roc(response = d.test.set1$Churn, predictor = as.numeric(pred4))
#plot(rf.roc, col = "red" , add = TRUE, print.auc.y = 0.6, print.auc = TRUE)
  
# plotting ROC curve
perf <- performance(pr, measure = "tpr", x.measure = "fpr")

par(mar=c(5,5,2,2),xaxs = "i",yaxs = "i",cex.axis=1.3,cex.lab=1.4)
# plotting the ROC curve
plot(perf,col="blue",lty=3, lwd=3)
# calculating AUC
auc <- performance(pr,"auc")
# now converting S4 class to vector
auc <- unlist(slot(auc, "y.values"))
# adding min and max ROC AUC to the center of the plot
minauc<-min(round(auc, digits = 2))
maxauc<-max(round(auc, digits = 2))
minauct <- paste(c("(AUC)  = "),minauc,sep="")
#maxauct <- paste(c("max(AUC) = "),maxauc,sep="")
legend(0.3,0.6,c(minauct,"\n"),border="white",cex=1.7,box.col = "red")
```

```{r}
library(ROCR)
library(pROC)
# need to create prediction object from ROCR
pr <- prediction(as.numeric(y_pred), as.numeric(d.test.set1$Churn))
#rf.roc <- roc(response = d.test.set1$Churn, predictor = as.numeric(pred4))
#plot(rf.roc, col = "red" , add = TRUE, print.auc.y = 0.6, print.auc = TRUE)
  
# plotting ROC curve
perf <- performance(pr, measure = "tpr", x.measure = "fpr")

par(mar=c(5,5,2,2),xaxs = "i",yaxs = "i",cex.axis=1.3,cex.lab=1.4)
# plotting the ROC curve
plot(perf,col="blue",lty=3, lwd=3)
# calculating AUC
auc <- performance(pr,"auc")
# now converting S4 class to vector
auc <- unlist(slot(auc, "y.values"))
# adding min and max ROC AUC to the center of the plot
minauc<-min(round(auc, digits = 2))
maxauc<-max(round(auc, digits = 2))
minauct <- paste(c("(AUC)  = "),minauc,sep="")
#maxauct <- paste(c("max(AUC) = "),maxauc,sep="")
legend(0.3,0.6,c(minauct,"\n"),border="white",cex=1.7,box.col = "red")
```

## SVM with linear kernel model with SMOTE sampling

```{r}
library(e1071)
set.seed(1)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
train.set1 = newdata[d.data1,] 
test.set1 = newdata[-d.data1,]
train.set1$Churn = as.factor(train.set1$Churn)
test.set1$Churn = as.factor(test.set1$Churn)
svc.fit <- svm(Churn ~ ., data = train.set1, kernel = "linear", cost = 0.01)
summary(svc.fit)



```
the training and test error rates? 
```{r}
train.pred <- predict(svc.fit,train.set1)
table(train.set1$Churn, train.pred)
```
```{r}

cat("Training error rate is",(549+729)/(1901+2054+549+729))
```

```{r}
test.pred <- predict(svc.fit, test.set1)
table(test.set1$Churn, test.pred)

```
```{r}

cat("Test error rate is",(237+297)/(811+898+237+297))
```

## LASSO model with SMOTE sampling
```{r}
set.seed(1)
library(glmnet)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
train.set1 = newdata[d.data1,] 
test.set1 = newdata[-d.data1,]
train.set1$Churn=as.numeric(train.set1$Churn)
test.set1$Churn=as.numeric(test.set1$Churn)
matrix.train.set1<- model.matrix(Churn ~ ., data = train.set1) 
matrix.test.set1 <- model.matrix(Churn ~ ., data = test.set1) 
grid <- 10 ^ seq(10, -2, length = 100) 
fit.lasso <- glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
set.seed(30)
cv.lasso <- cv.glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
bestlam.lasso <- cv.lasso$lambda.min
bestlam.lasso
pred.lasso <- predict(fit.lasso, s = bestlam.lasso, newx = matrix.test.set1)
mean((pred.lasso - test.set1$Churn)^2)

```

## Logistic regression model with SMOTE sampling

```{r}
library(MASS)
set.seed(1)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.80)) 
train.set1 = newdata[d.data1,] 
test.set1 = newdata[-d.data1,]
logistic.reg.model = glm(Churn ~., data = train.set1 , family="binomial")
summary(logistic.reg.model)
pred1 = predict(logistic.reg.model, test.set1, type = "response")
pred1.glm = rep(0, length(pred1))
pred1.glm[pred1 > 0.5] = 1
table(pred1.glm, test.set1$Churn)
meanval1 = mean(pred1.glm != test.set1$Churn)
meanval1

cat("Test error of the Logistic Regression model is", meanval1*100,"%")

```



# Performing EDA for different situations to give insights for final recommendation

```{r}
# Finding the median of totalcharges
median(cleanData$TotalCharges)
```

```{r}

# Condition -> People with totalcharges greater than median and opting/not opting for PaperlessBilling with different contracts

x = subset(d.cleanData, TotalCharges > 1400 & PaperlessBilling == "0" & Contract.One.year == "1")
x1= subset(x,Churn=="0")
x2= subset(x,Churn=="1")
a = subset(d.cleanData, TotalCharges > 1400 & PaperlessBilling == "0" & Contract.Two.year == "1")   
a1= subset(a,Churn=="0")
a2= subset(a,Churn=="1")
b = subset(d.cleanData, TotalCharges > 1400 & PaperlessBilling == "0" & Contract.Monthtomonth == "1")
b1= subset(b,Churn=="0")
b2= subset(b,Churn=="1")


x_new = subset(d.cleanData, TotalCharges > 1400 & PaperlessBilling == "1" & Contract.One.year == "1")
x1_new= subset(x_new,Churn=="0")
x2_new= subset(x_new,Churn=="1")
a_new = subset(d.cleanData, TotalCharges > 1400 & PaperlessBilling == "1" & Contract.Two.year == "1")   
a1_new= subset(a_new,Churn=="0")
a2_new= subset(a_new,Churn=="1")
b_new = subset(d.cleanData, TotalCharges > 1400 & PaperlessBilling == "1" & Contract.Monthtomonth == "1")
b1_new= subset(b_new,Churn=="0")
b2_new= subset(b_new,Churn=="1")


#x #346 (Total Number of people with TotalCharges > 1400 & PaperlessBilling == "0" & Contract.One.year == "1" )
#x1 #310 (Total number of people with above condition and churn=0)
#x2 #36 (Total number of people with above condition and churn=1)

#a #534 (Total Number of people with TotalCharges > 1400 & PaperlessBilling == "0" & Contract.Two.year == "1" )
#a1 #522 (Total number of people with above condition and churn=0)
#a2 #12 (Total number of people with above condition and churn=1)

#b #305 (Total Number of people with TotalCharges > 1400 & PaperlessBilling == "0" & Contract.Monthtomonth == "1" )
#b1 #228 (Total number of people with above condition and churn=0)
#b2 #77 (Total number of people with above condition and churn=1)

#x_new # 642 (Total Number of people with TotalCharges > 1400 & PaperlessBilling == "1" & Contract.One.year == "1" )
#x1_new #538 (Total number of people with above condition and churn=0)
#x2_new #104 (Total number of people with above condition and churn=1)

#a_new #662 (Total Number of people with TotalCharges > 1400 & PaperlessBilling == "1" & Contract.Two.year == "1" )
#a1_new #631 (Total number of people with above condition and churn=0)
#a2_new #31 (Total number of people with above condition and churn=1)

#b_new #1023 (Total Number of people with TotalCharges > 1400 & PaperlessBilling == "1" & Contract.Monthtomonth == "1" )
#b1_new #625 (Total number of people with above condition and churn=0)
#b2_new #400 (Total number of people with above condition and churn=1)


```

```{r}

counts=c(538,631,625)
barplot(counts, names.arg=c("Contract=1yr","Contract=2yr","Contract=Monthly"), main="Total charges > 1400 and No PaperlessBilling and Churn=0")
counts1=c(104,31,400)
barplot(counts1, names.arg=c("Contract=1yr","Contract=2yr","Contract=Monthly"), main="Total charges > 1400 and No PaperlessBilling and Churn=1")
```

```{r}
counts=c(310,522,228)
barplot(counts, names.arg=c("Contract=1yr","Contract=2yr","Contract=Monthly"), main="Total charges > 1400 and PaperlessBilling and Churn=0")
counts1=c(36,12,77)
barplot(counts1, names.arg=c("Contract=1yr","Contract=2yr","Contract=Monthly"), main="Total charges > 1400 and PaperlessBilling and Churn=1")

#full = lm(Churn~tenure+Contract+MonthlyCharges+PaperlessBilling+PaymentMethod+TotalCharges+InternetService,data=cleanData)
#null=lm(Churn ~1,data=cleanData)
#step(null,scope=list(lower=null, upper=full),direction="forward",data=cleanData)
#new1=step(lm(Churn~tenure+Contract+MonthlyCharges+PaperlessBilling+PaymentMethod+TotalCharges+InternetService,data=cleanData),direction="backward")
```


# Models using only two variables MonthlyCharges and Total Charges (LASSO, Boosting and Random Forest)


```{r}
library(MASS)
set.seed(10)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.80)) 
train.set1 = newdata[d.data1,] 
test.set1 = newdata[-d.data1,]
logistic.reg.model = glm(Churn ~ MonthlyCharges+TotalCharges, data = train.set1 , family="binomial")
summary(logistic.reg.model)
pred1 = predict(logistic.reg.model, test.set1, type = "response")
pred1.glm = rep(0, length(pred1))
pred1.glm[pred1 > 0.5] = 1
table(pred1.glm, test.set1$Churn)
meanval1 = mean(pred1.glm != test.set1$Churn)
meanval1

cat("Test error of the Logistic Regression model is", meanval1*100,"%")

```

```{r}
#LASSO
set.seed(222)
library(glmnet)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
train.set1 = newdata[d.data1,] 
test.set1 = newdata[-d.data1,]
train.set1$Churn=as.numeric(train.set1$Churn)
test.set1$Churn=as.numeric(test.set1$Churn)
matrix.train.set1<- model.matrix(Churn ~ MonthlyCharges+TotalCharges, data = train.set1) 
matrix.test.set1 <- model.matrix(Churn ~ MonthlyCharges+TotalCharges, data = test.set1) 
grid <- 10 ^ seq(10, -2, length = 100) 
fit.lasso <- glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
set.seed(30)
cv.lasso <- cv.glmnet(matrix.train.set1, train.set1$Churn, alpha = 1, lambda = grid, thresh = 1e-12)
bestlam.lasso <- cv.lasso$lambda.min
bestlam.lasso
pred.lasso <- predict(fit.lasso, s = bestlam.lasso, newx = matrix.test.set1)
mean((pred.lasso - test.set1$Churn)^2)
```

```{r}
library(gbm)
set.seed(1)
#Boosting
data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
train.set1 = newdata[data1,] 
test.set1 = newdata[-data1,]
#converting response variable to char type to be used in the boosting model
train.set1$Churn = as.character(train.set1$Churn)
test.set1$Churn = as.character(test.set1$Churn)
fit.boost=gbm(Churn∼MonthlyCharges+TotalCharges, data = train.set1, distribution = "bernoulli", n.trees=500, interaction.depth=4, shrinkage = 0.1)
pred4 = predict(fit.boost, newdata = test.set1, n.trees = 500)
pred4 = ifelse(pred4 > 0.5,1,0)
boost.table = table(pred4, test.set1$Churn)

accuracy = (boost.table[1,1] + boost.table[2,2]) / (dim(test.set1)[1])
cat("Accuracy of the Boosting Model is", accuracy*100)
```

```{r}
#Random forest
library(randomForest)
d.data1 = sort(sample(nrow(newdata), nrow(newdata)*.70)) 
d.train.set1 = newdata[d.data1,] 
d.test.set1 = newdata[-d.data1,]



levels(d.test.set1$MonthlyCharges)<-levels(d.train.set1$MonthlyCharges)
levels(d.test.set1$TotalCharges)<-levels(d.train.set1$TotalCharges)


set.seed(123)
y_test <- d.test.set1["Churn"]
fit1 <- randomForest(Churn ~TotalCharges+ MonthlyCharges , data=d.train.set1, mtry = 26, nTree=500, importance=TRUE)
fit1
y_pred=predict(fit1,newdata = d.test.set1)

#Using IMportance to find out which variables are important
importance(fit1)

t1=table(y_pred, d.test.set1$Churn)
Accuracy = (t1[1,1] + t1[2,2]) / (dim(d.test.set1)[1])
cat("Accuracy is ", Accuracy)
#plot(fit1)
```



